---
title: "Statistical Analyses"
author: "David John Baker"
date: "21/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This markdown contains the analysis for "The Attraction of Predominant Chords".
It attempts to more clearly link the theoretical introduction to the statistical modeling to the discussion.

The document is meant to serve as a reference to be pasted into the manuscript.
All code to create tables, charts, and statistical analyses are found here.

## Hypotheses and Analyses 

### Supervised 

1. Top Down Model: We predict category of predominant will predict attraction ratings globally.
  - Chromatic most attracted
  - Diatonic second
  - Bridge third
  - Non PD last 
  - **Follow up with pairwise comparisons**
2. We predict musical training will lead to qualitative difference in ratings (interaction/random slopes)
3. We predict feature analysis will help structure what it is about chords that leads to differences in ratings

### Unsupervised 

Following our regressionthe modeling of these five musical features, a final analysis using unsupervised machine learning methods to explore participant response strategies(James et al. 2013). Mirroring our above strategy to take both a bottom up and top down approach to our perceptual data using supervised methods, we also employ a top down and bottom up analysis using unsupervised methods. 

We first decompose the empirical response data from our experiment using principal component analysis (PCA); this analysis serves as our top down method of understanding listener response strategies. 

Secondly, we submit the features used in our feature-based regression model to a hierarchical agglomerative clustering algorithm to consider novel ways the pre-dominant category might be re-conceptualized. 

## Data Import 

```{r, include=FALSE}
library(readr)
library(stringr)
library(ggplot2)
library(tidyr)
library(GGally)
library(MuMIn)
library(janitor)
library(lme4)
library(lmerTest)
library(patchwork)
library(gt)
library(psych)
library(Hmisc)
library(forcats)
library(multcomp)
library(ggpubr)
library(dplyr)
library(sjPlot)
library(sjlabelled)
library(forcats)
library(janitor)
library(tidymodels)

# Importing Data
## Experimental Data 
response_data <- read_csv("tidy_tables/response_data.csv")

chord_stimuli_data <- read_csv("tidy_tables/chord_experiment_stimuli_data.csv")

chord_stimuli_data <- chord_stimuli_data %>% 
  rename(chord = chord_symbol,
         chord_family = group) 

# Chord Features
chord_features <- read_csv("tidy_tables/chord_features.csv")
chord_features <- janitor::clean_names(chord_features)

chord_features <- chord_features %>%
  separate(col = stimuli, into = c("chord_id","temp"), "\\.") %>%
  separate(col = temp, into = c("chord","notes"), sep = "\\:") %>%
  mutate(chord = str_trim(chord)) %>%
  mutate(notes = str_trim(notes))

# Demographic Data
gmsi <- read_csv("tidy_tables/gmsi_table.csv")
gmsi <- janitor::clean_names(gmsi)
gmsi <- gmsi %>% 
  rename(experimental_group = group,
         participant = participant_no)

# Build Big Data 
df <- response_data %>%
  pivot_longer(cols = I:`viio7/V`,names_to = "chord",values_to = "rating") %>%
  select(participant, chord, rating, block, group, `How attracted`) %>%
  left_join(chord_stimuli_data) %>%
  left_join(gmsi) %>%
  left_join(chord_features)
```

## Descriptives

```{r}
df %>%
  select(participant, group, i_have_had_formal_training_in_music_theory_for_years) %>%
  distinct() %>%
  group_by(group) %>%
  summarise(mean_theory = mean(i_have_had_formal_training_in_music_theory_for_years),
            sd_theory = sd(i_have_had_formal_training_in_music_theory_for_years))

```

## Results

### Supervised 

First we check for test-restest reliability then collapse data based on R1F  =  0.92 Reliability of average of all items for one  time (Random time effects) and RkF  =  0.96 Reliability of average of all items and both times (Fixed time effects).

```{r}
# Check test-retest with psych package 
response_data %>%
  select(block,I:`viio7/V`) %>%
  rename(time = block) %>%
  data.frame() %>%
  testRetest()

# Reduce df to variables for modeling 
model_data <- df %>%
  select(rating, 
         chord,
         chord_family,
         lerdhal_tension,
         parncut_roughness,
         semitone_voice_move, 
         rootmotion,
         number_tendency_tones4_6,
         participant,
         block, 
         i_have_had_formal_training_in_music_theory_for_years, 
         f3_musical_training,experimental_group) %>%
  mutate(experimental_group = factor(experimental_group, c("prolific", "freshman","upperclass"))) %>%
  mutate(chord_family_f = factor("non_pd","bridge","common_pd","chromatic"))

# Test retest, OK, collapse to one average rating per participant
model_data %>%
  group_by(chord, participant) %>%
  summarise(mean_rating = mean(rating)) -> summary_ratings

model_data %>%
  select(chord, chord_family, lerdhal_tension:participant, i_have_had_formal_training_in_music_theory_for_years, f3_musical_training, experimental_group) %>%
  left_join(summary_ratings) -> collapsed_model
```

## Regression Analyses

We now run each hypothesis as linear, mixed effects regression model.
We note here that [ANOVA models are a special case of regression](https://lindeloev.github.io/tests-as-linear/) and that running a linear mixed effects model with a random effect of participant is [analgous to running a within subjects repeated measures ANOVA in R](https://m-clark.github.io/docs/mixedModels/anovamixed.html).


### Chord Category Model 

```{r}
# CHORD CATEGORY
# RM-ANOVA as LINEAR MIXED EFFECTS
chord_cat_rm <- lmer(mean_rating ~ chord_family + (1|participant), data = collapsed_model)

# Model Summary Output 
summary(chord_cat_rm)
anova(chord_cat_rm)
r.squaredGLMM(chord_cat_rm)

# Post Hoc Comparisions 
glht(chord_cat_rm, mcp(chord_family="Tukey")) %>% summary()
```

#### Figure For Chord Category Model 

```{r}
means_a <- tibble(means = c(4.39518,4.39518+.41810,4.39518+.52897,4.39518+.03238),
                std_errors = c(.09757,.05051,.05051,.05229),
                labz = c("Bridge Chords","Chromatic Predominant","Diatonic Predominant","Non Predominant"))


means_a %>%
  mutate(labz = factor(labz, levels = c("Bridge Chords","Chromatic Predominant","Diatonic Predominant","Non Predominant"))) %>%
  ggplot(aes(y = means, x = labz)) +
  geom_bar(stat = "identity") +
  geom_bracket(
    xmin = c("Bridge Chords","Bridge Chords", "Non Predominant","Non Predominant"), 
    xmax = c("Chromatic Predominant","Diatonic Predominant","Chromatic Predominant","Diatonic Predominant"),
    y.position = c(6, 7, 8 , 6), label = c("***"),
    tip.length = 0.1) +
#  geom_errorbar(aes(ymin = means - std_errors, ymax = means+std_errors)) +
  labs(x = "Chord Category", y = "Adjusted Means", title = "Chord Category Model") +
  theme_minimal() -> analysis_a_plot_1

analysis_a_plot_1

# Same analysis, run with base R calculations 
#chord_cat_rm_anova <- aov(mean_rating ~ chord_family + Error(participant), data = collapsed_model)
#summary(chord_cat_rm_anova)
```

### Musical Training Model 

```{r}
# MUSICAL TRAINING
# Run as Linear Mixed Effects
# Main Effects: Theory Training and Chord Category
# Interaction 
# Participant as Random Effects, within subject 


musical_training_rm <- lmer(mean_rating ~ chord_family*i_have_had_formal_training_in_music_theory_for_years  + (1|participant), 
                            data = collapsed_model)
summary(musical_training_rm)

anova(musical_training_rm)
r.squaredGLMM(musical_training_rm)

# Also run with base-R 
# rm_anova_music <- aov(mean_rating ~ chord_family*i_have_had_formal_training_in_music_theory_for_years + Error(participant), data = collapsed_model)
# summary(rm_anova_music)
```

#### Figure for Musical Training Model 

```{r}
library(ggeffects)

vexta = c("0","0.5","1","2","3","4 - 6","> 7")

ggpredict(musical_training_rm, 
          terms = c("i_have_had_formal_training_in_music_theory_for_years", "chord_family"), 
          type = "fe") %>% 
   plot(ci.style = "ribbon", colors = "bw", alpha = .05) +
   labs(x = "I have had formal training in music theory for __ years", y = "Rating", 
        title = "Music Theory Training Model") +
   scale_x_continuous(breaks = seq(1,7,1), 
                     labels = vexta)  -> rm_musical_plot
   

rm_musical_plot$labels$linetype <- "Chord Category"

rm_musical_plot$data$group <- rm_musical_plot$data$group %>% str_replace_all("bridge_chords", "Bridge Chords")
rm_musical_plot$data$group <- rm_musical_plot$data$group %>% str_replace_all("not_pd", "Non predominant")
rm_musical_plot$data$group <- rm_musical_plot$data$group %>% str_replace_all("common_pd", "Diatonic Predominant")
rm_musical_plot$data$group <- rm_musical_plot$data$group %>% str_replace_all("chromatic_pd", "Chromatic Predominant")

rm_musical_plot

# Descriptive Panel 
model_data %>%
  mutate(experimental_group = as.factor(str_to_title(experimental_group))) %>%
  select(experimental_group, i_have_had_formal_training_in_music_theory_for_years, participant) %>%
  distinct() %>%
  ggplot(aes(
    x = fct_reorder(experimental_group, i_have_had_formal_training_in_music_theory_for_years), 
    y = i_have_had_formal_training_in_music_theory_for_years)) +
  geom_jitter() +
  scale_y_continuous(breaks = seq(1,7,1), 
                     labels = vexta) +
  stat_summary(fun = "mean", colour = "red", size = 2, geom = "point") +
  stat_summary(fun.data = "mean_cl_boot", colour = "black", size = 1) +
  coord_flip() +
  labs(title = "Distribution of Music Theory Training",
       subtitle = "Question 38 from Goldsmiths Musical Sophistication Index",
       x = "Participant Group",
       y = "I have had formal training in music theory for __ years") +
  theme_minimal() -> descriptive_theory_plot

descriptive_theory_plot$data$experimental_group <- descriptive_theory_plot$data$experimental_group %>% str_replace_all("Upperclass", "Junior/Senior")

descriptive_theory_plot
#---------------------------------------------------------------------


```

### Music Feature + Training Model 

```{r}
all_musical_indv_theory_model <- lmer(scale(rating) ~  scale(lerdhal_tension) + 
                                      scale(parncut_roughness) + scale(semitone_voice_move) + 
                                        as.character(rootmotion) + scale(number_tendency_tones4_6) +
                                        scale(i_have_had_formal_training_in_music_theory_for_years) +
                        (1+scale( i_have_had_formal_training_in_music_theory_for_years)|participant), 
                                    data = model_data)


r.squaredGLMM(all_musical_indv_theory_model)

```

#### Figure for Musical Training Model 

```{r}

(re.effects <- plot_model(all_musical_indv_theory_model, type = "est", show.values = TRUE,colors = c("black","black")))

levels(re.effects$data$term) <- c("Music Theory Training", "^4, ^#4, b^6, ^6" , "Root Motion",
                                  "Semitonal Voice Leading","Chord Roughness","Chord Distance")

re.effects + 
  labs(title = "Musical Feature and Musical Training Model", subtitle = "Fixed Effects") + 
  theme_minimal() -> model_3_figure 

model_3_figure

ggsave(plot = model_3_figure, 
       filename = "img/Figures/Figure_Model_Three_fixed_effects_.png",
              height = 6, width = 9, units = "in", dpi = 300)
```

#### Combine to One Panel 

```{r}
(analysis_a_plot_1 + descriptive_theory_plot )/ (rm_musical_plot + model_3_figure) -> compound_plot

compound_plot

ggsave(filename = "img/Figures/AnalysisPanel.png", height = 9, width = 13, units = "in", dpi = 300)

# ALL MODELS HERE 
#tab_model(chord_cat_rm, musical_training_rm,feature_model2)
#anova(chord_cat_rm, musical_training_rm)
```


## Unsupervised Analyses 

The next two sections show code used for the PCA and HCA.

```{r}
# Re-Import Tables
# Get only rating data
attraction_ratings <- response_data %>%
  dplyr::select(I:`viio7/V`, group, participant)

# Demographic Data
gmsi <- read_csv("tidy_tables/gmsi_table.csv")
gmsi <- janitor::clean_names(gmsi)
gmsi <- gmsi %>% 
  rename(experimental_group = group,
         participant = participant_no)

gmsi %>%
  dplyr::select(participant, i_have_had_formal_training_in_music_theory_for_years) -> gmsi_2

attraction_ratings <- attraction_ratings %>% 
  left_join(gmsi_2)
  
chord_stimuli_data <- read_csv("tidy_tables/chord_experiment_stimuli_data.csv")

chord_stimuli_data <- chord_stimuli_data %>% 
  rename(chord = chord_symbol,
         chord_family = group) 
```

The object `attraction_ratings` now has all the ratings of the chords and the group identifier from where the rating came from. 

Next, we prep and run the PCA on this data.
Experimental group is saved to be an identifier.

```{r}
# Make Exploratory Plot 
pca_rec <- recipe(~., data = attraction_ratings ) %>%
    update_role(group, participant, i_have_had_formal_training_in_music_theory_for_years, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

pca_prep <- prep(pca_rec)

pca_prep

tidied_pca <- tidy(pca_prep, 2)

chord_stimuli_data_table <- chord_stimuli_data %>%
  rename(terms = chord)

tidied_pca %>%
  left_join(chord_stimuli_data_table) -> tidied_pca2
```

Now we have scores for every single chord on every single dimension. 

Want to know if our PCA was helpful at all first. 

* First PC explains ~25% of variation
* Next two are also formidable

```{r}
# Extract PCA Variance 
sdev <- pca_prep$steps[[2]]$res$sdev

percent_variation <- sdev^2 / sum(sdev^2)

sum(percent_variation[1:3])
sum(percent_variation[1])
sum(percent_variation[2])
sum(percent_variation[3])

tibble(
  component = unique(tidied_pca$component),
  percent_var = percent_variation ## use cumsum() to find cumulative, if you prefer
) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(component, percent_var)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL, y = "Percent Variance",
       title = "Cumulative Variance Explained") +
  theme_minimal() +
  coord_flip() -> var_explained_plot

var_explained_plot

ggsave(filename = "img/Figures/Figure_Variation_Explained.png", 
       height = 6, width = 9, dpi = 300, plot = var_explained_plot)

```

Next we plot the first five principle components with all items.

Main take aways from plot

* PC1 appears to be just general attraction rating
* All scores positive reflect the fact that all asked to do same task
* Going to look at each in depth after

```{r}
# PLOTS HERE NOT IN MANUSCRIPT 
# EXCLUDE 
tidied_pca2 %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = chord_family)) +
  geom_col() +
  facet_wrap(~component, nrow = 1) +
  labs(y = "Chord", x = "Loading", title = "First Five Principal Components") -> pc15

pc15

ggsave(filename = "img/Figures/Figure_PC1to5.png", height = 6, width = 9, dpi = 300, plot = pc15)

```

* PC1 is Pre Dominant Attraction  

```{r}
tidied_pca2 %>%
  filter(component %in% paste0("PC", 1)) %>%
  mutate(component = fct_inorder(component)) %>%
  mutate(value = round(value, 3)) %>%
  mutate(chord_family = str_replace_all(chord_family, "_"," ")) %>%
  mutate(chord_family = str_to_title(chord_family)) %>%
  mutate(chord_family = str_replace_all(chord_family,"Pd","PDs")) %>%
  ggplot(aes(x = value, y = reorder(terms, abs(value)))) +
  geom_point(aes(shape = chord_family), size = 3) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(y = "Chord", x = "Loading",
       title = "Response Strategy: Experimental Prompt", shape = "Chord Category") -> pc1


#   scale_fill_manual(labels = c("Bridge Chords","Chromatic PDs","Common PDs","Non PDs"), values = c("green","black","blue","red")) +
pc1

ggsave(filename = "img/Figures/PC1.png", 
       height = 9, width = 5, dpi = 300, plot = pc1)

```

* PC2 is sensitivy to training, more evident through next chart

```{r}
tidied_pca2 %>%
  filter(component %in% paste0("PC", 2)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(x = value, y = reorder(terms, abs(value)), fill = chord_family)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(labels = c("Bridge Chords","Chromatic PDs","Common PDs","Non PDs"),
                    values = c("green","black","blue","red")) +
  labs(y = "Chord", title = "PC2: Sensitivity to Training", fill = "Chord Family") -> pc2

pc2

ggsave(filename = "img/Figures/PC2.png", height = 6, width = 9, dpi = 300, plot = pc2)
```

Here we use loadings on each rating, plot PC1 vs PC2

```{r}
# PANEL 2 
juice(pca_prep) %>%
  mutate(group = str_replace_all(group, "upperclass","Junior/Senior")) %>%
  mutate(group = str_to_title(group)) %>%
  mutate(`Experimental Group` = str_to_title(group)) %>%
  ggplot(aes(PC1, PC2, 
             shape = `Experimental Group`)) +
  geom_point(alpha = 1) +
  scale_shape_manual(values=c( 8, 13, 4))+
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(title = "Scores of Items from PCA Loadings",
       x = "PC1: Predominant Attraction",
       y = "PC2: Sensitivity to Music Theory Training",
       shape = "Participant Group") -> pc1vpc2

pc1vpc2

ggsave(filename = "img/Figures/PC1vsPC2.png", height = 6, width = 9, dpi = 300, plot = pc1vpc2)
```

#### Panel 3 for Manuscript

```{r}
# Make Panel
library(patchwork)
(pc1 + (pc1vpc2)) -> pca_panel
ggsave(plot = pca_panel, filename = "img/Figures/PCA_Panel.png", height = 9, width = 12, dpi = 300)
```


```{r}
# Old Exploratory Plots, Not in Manuscript 
juice(pca_prep) %>%
  ggplot(aes(PC1, PC3, color = group)) +
  geom_point(alpha = 0.7, size = 2) +
  labs() -> pc1vpc3

ggsave(filename = "img/unsupervised/pc1vpc3.png", height = 6, width = 9, dpi = 300, plot = pc1vpc3)

juice(pca_prep) %>%
  ggplot(aes(PC2, PC3, color = group)) +
  geom_point(alpha = 0.7, size = 2) +
  labs() -> pc2vpc3

ggsave(filename = "img/unsupervised/pc2vpc3.png", height = 6, width = 9, dpi = 300, plot = pc2vpc3)
```

More Exploration, not included 

```{r}
# Correlation with general average 
# tidied_pca2 %>%
#   filter(component == "PC1") %>%
#   dplyr::select(terms, value) %>%
#   rename(chord = terms) -> pca_terms
# 
# 
# feature_model_data %>%
#   dplyr::select(mean_chord_ratings, chord) %>%
#   left_join(pca_terms)-> cor_pca
# 
# cor_pca %>%
#   ggplot(aes(x = value, y = mean_chord_ratings, label = chord)) +
#   geom_point() +
#   geom_text(nudge_y = .1) +
#   geom_smooth()
# 
# cor.test(cor_pca$mean_chord_ratings, cor_pca$value)
```

### HCA 

```{r}

chord_stimuli_data <- read_csv("tidy_tables/chord_experiment_stimuli_data.csv")

chord_stimuli_data <- chord_stimuli_data %>% 
  rename(chord = chord_symbol,
         chord_family = group) 

# Chord Features
chord_features <- read_csv("tidy_tables/chord_features.csv")
chord_features <- janitor::clean_names(chord_features)

chord_features <- chord_features %>%
  separate(col = stimuli, into = c("chord_id","temp"), "\\.") %>%
  separate(col = temp, into = c("chord","notes"), sep = "\\:") %>%
  mutate(chord = str_trim(chord)) %>%
  mutate(notes = str_trim(notes))

chord_features %>%
  select(parncut_roughness:number_4_b6) -> cluster_features

rownames(cluster_features) <- chord_features$chord


cluster_features %>% scale() -> scaled_cluster_features

dist_cluster <- dist(scaled_cluster_features)

dist_cluster_object <- hclust(dist_cluster, method = "complete")

par(mfrow=c(1,1))

jpeg("img/Figures/Figure4.jpeg", height = 500, width = 900)

plot(dist_cluster_object , xlab="", sub="", cex=1.5, main = "HCA of Chord Features")
dev.off()

# Linage does not seem to affect too much in terms of general categories 
dist_cluster_object_complete <- hclust(dist_cluster,method = "complete")
dist_cluster_object_average <- hclust(dist_cluster,method = "average")
dist_cluster_object_single <- hclust(dist_cluster, method = "single")

par(mfrow=c(1,3))

plot(dist_cluster_object_complete ,main="Complete Linkage", xlab="", sub="",
cex=.9)

plot(dist_cluster_object_average , main="Average Linkage", xlab="", sub="",
cex=.9)

plot(dist_cluster_object_single, main="Single Linkage", xlab="", sub="",
cex=.9)

```

